---
title: "Getting started with rmzTab-M"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting started with rmzTab-M}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(rmzTabM)
library(kableExtra)
```

## Reading mzTab-M files
You can read an mzTab-M file in *tab separated format* into a data frame structure as follows:
```{r readMzTabM1, warning=FALSE}
testfile <- system.file("testdata", c("lipidomics-example.mzTab"), package="rmzTabM")
mzTabTable <- readMzTab(testfile)
kable(head(mzTabTable[,1:3])) %>% kable_styling(bootstrap_options = c("striped", "hover", "condensed", font_size = 7)) %>% scroll_box(width = "800px", height = "200px")
```

You can extract the individual section tables from this one as follows:

```{r mzTabMExtract, warning=FALSE}
mtdTable <- extractMetadata(mzTabTable)
smlTable <- extractSmallMoleculeSummary(mzTabTable)
smfTable <- extractSmallMoleculeFeatures(mzTabTable)
smeTable <- extractSmallMoleculeEvidence(mzTabTable)

knitr::kable(head(smlTable)) %>% kable_styling(bootstrap_options = c("striped", "hover", "condensed", font_size = 7)) %>% scroll_box(width = "800px", height = "200px")

knitr::kable(head(smfTable)) %>% kable_styling(bootstrap_options = c("striped", "hover", "condensed", font_size = 7)) %>% scroll_box(width = "800px", height = "200px")

knitr::kable(head(smeTable)) %>% kable_styling(bootstrap_options = c("striped", "hover", "condensed", font_size = 7)) %>% scroll_box(width = "800px", height = "200px")
```


To turn these tables into objects, use the R6 class constructor method `new()`:

```{r readMzTabMObj, warning=FALSE}
mzTabObject <- MzTab$new()$fromDataFrame(mzTabTable)
```

You can then access sections like Metadata as object members:

```{r readMzTabMMetadata, warning=FALSE}
# this is an R6 object
metadata <- mzTabObject$metadata
# these are lists
smallMoleculeSummaryEntries <- mzTabObject$smallMoleculeSummary
smallMoleculeFeatureEntries <- mzTabObject$smallMoleculeFeature
smallMoleculeEvidenceEntries <- mzTabObject$smallMoleculeEvidence
```

Extracting values is possible from either representation, depending on whether you prefer a tabular style or an object oriented style, however, there may be type differences:

```{r extractValues, warning=FALSE}
# this is the SmallMoleculeSummary list first entry id
smallMoleculeSummaryEntries[[1]]$sml_id

# and this is the same with the data frame
as.numeric(smlTable$SML_ID)
```

## Writing mzTab-M files

If you have an mzTab-M data frame, you can write it as follows:

```{r writeMzTabMTable1, warning=FALSE}
utils::write.table(
    mzTabTable,
    file = file.path(tempdir(check=TRUE), "mzTabWrite1.mztab"),
    row.names = FALSE,
    col.names = FALSE,
    quote = FALSE,
    sep = "\t",
    na = "",
    fileEncoding = "UTF8"
  )
```

For an MzTab object, you can write to the tab separated format:

```{r writeMzTabMTable2, warning=FALSE}
writeMzTab(mzTabObject, file.path(tempdir(check=TRUE), "mzTabWrite2.mztab"))
```

Or to JSON format:
```{r writeMzTabMTable3, warning=FALSE}
writeMzTabJSON(mzTabObject, file.path(tempdir(check=TRUE), "mzTabWrite3.mztab.json"))
```

## Validating mzTab-M files

To validate an mzTab-M file, you can access the mzTab Validator web application at https://apps.lifs-tools.org/mztabvalidator
<!-- rmzTab-M provides an API client to perform the validation against the web application: -->

<!-- ```{r validate, warning=FALSE} -->
<!--   validationMessages1 <- validateMzTab( -->
<!--     mzTabObject, -->
<!--     validationMode = "json", -->
<!--     validationLevel = "info", -->
<!--     maxErrors = 100, -->
<!--     semanticValidation = FALSE -->
<!--   ) -->
<!--   validationMessages1 -->
<!-- ``` -->

You can set the `validationLevel` to one of `info`, `warning` or `error`. If you enable `semanticValidation`, CV parameters present in your file will be checked against the [default recommended mapping file](https://github.com/HUPO-PSI/mzTab/blob/master/specification_document-releases/2_0-Metabolomics-Release/mzTab_2_0-M_mapping.xml).

In order to validate an mzTab-M file without needing to parse it locally, use the following call, which should return an empty list:

```{r validate1, warning=FALSE}
  validatePlainFile <- system.file("testdata", c("lipidomics-example.mzTab"),package="rmzTabM")
  mzTabString <- readChar(validatePlainFile, file.info(testfile)$size)
  validationMessages2 <- validateMzTab(
    mzTabString,
    validationMode = "plain",
    validationLevel = "info",
    maxErrors = 100,
    semanticValidation = FALSE
  )
  if(length(validationMessages2)==0) {
    print("No validation messages")
  } else {
    validationMessages2
  }
```
Alternatively, to run the validation with semantic checks of the used CV parameters against the default mapping file, which will give you hints on how to improve your file:

```{r validate2, warning=FALSE}
  validationMessages2 <- validateMzTab(
    mzTabString,
    validationMode = "plain",
    validationLevel = "info",
    maxErrors = 100,
    semanticValidation = TRUE
  )
  dfList<-lapply(validationMessages2, function(x) { data.frame("Category"=x$category, "Message"=x$code) })
  vmdf <- do.call("rbind", dfList)
  knitr::kable(vmdf) %>% kable_styling(bootstrap_options = c("striped", "hover", "condensed", font_size = 7)) %>% scroll_box(width = "800px", height = "1000px")
```
