---
title: "Conversion of an mzTab-M to MetaboLights MAF"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Conversion of an mzTab-M to MetaboLights MAF}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
  %\VignetteDepends{rmzTabM,metabolighteR,BiocStyle}
  %\VignettePackage{rmzTabM}
  %\VignetteKeywords{mass spectrometry, metabolomics, lipidomics}

---
```{r biocstyle, echo = FALSE, results = "asis"}
BiocStyle::markdown()
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

**Package**: `r Biocpkg("rmzTabM")`<br />
**Authors**: Steffen Neumann<br />
**Modified**: `r file.info("mzTabM2MAF.Rmd")$mtime`<br />
**Compiled**: `r date()`


```{r setup}
library(BiocStyle)
library(rmzTabM)
library(metabolighteR)
library(kableExtra)
```

# Converting an mzTab-M file to MetaboLights MAF 

The Metabolites Asignement File (MAF) is used by the EMBL-EBI MetaboLights 
repository to capture the metabolite abundances in the individual samples, 
the spectral characteristics of the features and, if present, the metabolite 
identifications. 

The `convertMzTab2MAF()` function will read the mzTab-M via `readMzTab()` 
and extract the `smlTable` and the `smfTable`. Abundances are extracted 
from the `smlTable`, while *m/z* and retention time come from the feature table. 

The MAF is created as a `data.frame` and written as TSV using `write.MAF()` 
from the `metabolighteR` package by Tom Wilson. The data.frame is also returned 
as (invisible) return value of `convertMzTab2MAF()`.

```{r convertMzTabM, warning=FALSE}

mztabfile <- system.file("testdata", 
                         c("lcmsms_dda_hydrophilic_height_mzTab.mztab"), 
                         package="rmzTabM")
MAFfile <- file.path(tempdir(check=TRUE), "m_MTBLS0815_v2_maf.tsv")                      

maf <- convertMzTab2MAF(mzTabfile=mztabfile,
                        MAFfile=MAFfile)

```



```{r analyseMAF, warning=FALSE}

kable(head(maf[!(grepl ("MSP: Unknown", maf$database_identifier) | is.na(maf$database_identifier)), ])) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", font_size = 7)) %>% 
  scroll_box(width = "800px", height = "200px")

```

# Session information

```{r sessionInfo}
sessionInfo()
```

# References
